Imaging of cortical oxygen tension and blood flow following targeted photothrombotic stroke

### 2018-06-24 article.tex
Re-added line numbers
Updated with Sergei's comments

% ----------------------------------------------------------------------------------------
% BEGIN ABSTRACT
% ----------------------------------------------------------------------------------------
\begin{abstract}
We present a dual-modality imaging system combining laser speckle contrast imaging and phosphorescence lifetime quenching to simultaneously map cortical blood flow and oxygen tension (\pO) in mice. Phosphorescence signal localization is achieved through the use of a digital micromirror device (DMD) that allows for selective excitation of arbitrary regions of interest. By targeting both excitation maxima of the oxygen-sensitive porphyrin Oxyphor PtG4, we are able to examine the effects of excitation wavelength on phosphorescence lifetime. We demonstrate the ability to measure differences in \pO\ between arteries and veins and large changes during an hyperoxic challenge. We dynamically monitor blood flow and \pO\ during DMD-targeted photothrombotic occlusion of an arteriole and highlight the presence of an ischemia-induced depolarization. Chronic tracking of the ischemic lesion over eight days revealed a rapid recovery, with the targeted vessel fully reperfusing and \pO\ returning to baseline values within five days. This system has broad applications for studying the acute and chronic pathophysiology of ischemic stroke and other vascular diseases of the brain.
\end{abstract}

@@

In this paper we present an update to the system by Ponticorvo and Dunn \cite{Ponticorvo:2010uv} that utilizes a more robust porphyrin probe \cite{Esipova:2011hi} and offers higher spatiotemporal resolutions for both LSCI and phosphorescence lifetime imaging. Using both of the phosphorescent probe's excitation maxima, we highlight wavelength-dependent differences in measured decay lifetimes. We demonstrate the ability to detect static variations in \pO\ between different types of vasculature and the dynamic monitoring of blood flow and \pO\ during DMD-targeted photothrombotic stroke. We also demonstrate chronic imaging by tracking the response to an ischemic event over several days. This system allows for acute and chronic imaging of relative blood flow and oxygen tension and has broad applications for both basic neuroscience and neuropathophysiology.

@@

\subsection{Oxygen Tension Measurements}
Oxyphor PtG4 is an oxygen-sensitive dendritic probe that contains Platinum(II)-meso-tetra-(3,5-dicarboxyphenyl)tetrabenzoporphyrin as the phosphorescent core \cite{Lebedev:2009cf,Esipova:2011hi} and has been effectively used to measure absolute oxygen tension using phosphorescence quenching \cite{Zhang:2013cd,Holt:2014cy}. It has two excitation maxima near 435 nm and 623 nm and an emission maximum at 782 nm (Fig.~\ref{fig:oxyphor_ptg4}a). Unlike previous generations of the Oxyphor probe, PtG4 is highly soluble in an aqueous environment without requiring the presence of environmental albumin for stabilization \cite{Esipova:2011hi}. For \textit{in vivo} measurements, Oxyphor PtG4 was introduced systemically via retro-orbital injection into the venous sinus for a target blood plasma concentration of 5 $\mu$M. Phosphorescence lifetime measurements were made in the time domain using the structured pulsed excitation light paradigm described above. The instrument response was accounted for by introducing a 2 $\mu$s temporal offset to the phosphorescent signal [$I(t)$] prior to fitting for the decay lifetime ($\tau$) in Eq.~(\ref{eq:phosphorescence}).

@@

The combination of laser speckle contrast imaging and phosphorescence lifetime imaging is a purely optical, non-contact strategy for measuring blood flow and oxygen tension within the cortex. This system builds upon prior work that used structured illumination to overcome traditional limitations of lifetime imaging \cite{Ponticorvo:2010uv}. Spatially patterning excitation light with a DMD allows for the use of a point detector, which offers both high sensitivity and speed for the collection of the phosphorescent signal. The spatial and temporal resolutions of the system are ultimately limited by noise. Regions as small as 0.01 mm$^{2}$ can reliably be excited at a pattern repetition rate of 10 Hz with 200 phosphorescent decays collected and averaged per pattern. Because intensity of the excitation light scales with the size of the target area, larger regions could allow for faster pattern rates at the expense of averaging (e.g. 100 Hz with 20-decay averaging). However, a 10 Hz pattern rate is more than sufficient for visualizing dynamic physiological events such as spreading depolarizations, which have been reported to propagate at a rate of several millimeters per minute \cite{Lauritzen:1994vs}.

@@

\acknowledgments
This study was supported by the National Institutes of Health (EB011556, NS078791, NS082518) and the American Heart Association (14EIA8970041).

>>>>

% ----------------------------------------------------------------------------------------
% BEGIN ABSTRACT
% ----------------------------------------------------------------------------------------
\begin{abstract}
We present a dual-modality imaging system combining laser speckle contrast imaging and oxygen-dependent quenching of phosphorescence to simultaneously map cortical blood flow and oxygen tension (\pO) in mice. Phosphorescence signal localization is achieved through the use of a digital micromirror device (DMD) that allows for selective excitation of arbitrary regions of interest. By targeting both excitation maxima of the oxygen-sensitive Oxyphor PtG4, we are able to examine the effects of excitation wavelength on the measured phosphorescence lifetime. We demonstrate the ability to measure differences in \pO\ between arteries and veins and large changes during an hyperoxic challenge. We dynamically monitor blood flow and \pO\ during DMD-targeted photothrombotic occlusion of an arteriole and highlight the presence of an ischemia-induced depolarization. Chronic tracking of the ischemic lesion over eight days revealed a rapid recovery, with the targeted vessel fully reperfusing and \pO\ returning to baseline values within five days. This system has broad applications for studying the acute and chronic pathophysiology of ischemic stroke and other vascular diseases of the brain.
\end{abstract}

@@

In this paper we present an update to the system by Ponticorvo and Dunn \cite{Ponticorvo:2010uv} that utilizes a more robust porphyrin probe \cite{Esipova:2011hi} and offers higher spatiotemporal resolutions for both LSCI and oxygen-dependent quenching of phosphorescence. Using both of the phosphorescent probe's excitation maxima, we highlight wavelength-dependent differences in measured decay lifetimes. We demonstrate the ability to detect static variations in \pO\ between different types of vasculature and the dynamic monitoring of blood flow and \pO\ during DMD-targeted photothrombotic stroke. We also demonstrate chronic imaging by tracking the response to an ischemic event over several days. This system allows for acute and chronic imaging of relative blood flow and oxygen tension and has broad applications for both basic neuroscience and neuropathophysiology.

@@

\subsection{Oxygen Tension Measurements}
Oxyphor PtG4 is an oxygen-sensitive dendritic probe that contains Platinum(II)-meso-tetra-(3,5-dicarboxyphenyl)tetrabenzoporphyrin as the phosphorescent core \cite{Lebedev:2009cf,Esipova:2011hi} and has been effectively used to measure absolute oxygen tension using phosphorescence quenching \cite{Zhang:2013cd,Holt:2014cy}. It has two excitation maxima near 435 nm and 623 nm and an emission maximum at 782 nm (Fig.~\ref{fig:oxyphor_ptg4}a). Oxyphor PtG4 is highly soluble in an aqueous environment without requiring the presence of environmental albumin for stabilization \cite{Esipova:2011hi}. For \textit{in vivo} measurements, Oxyphor PtG4 was introduced systemically via retro-orbital injection into the venous sinus for a target blood plasma concentration of 5 $\mu$M. Phosphorescence lifetime measurements were made in the time domain using the structured pulsed excitation light paradigm described above. The instrument response was accounted for by introducing a 2 $\mu$s temporal offset to the phosphorescent signal [$I(t)$] prior to fitting for the decay lifetime ($\tau$) in Eq.~(\ref{eq:phosphorescence}).

@@

The combination of laser speckle contrast imaging and oxygen-dependent quenching of phosphorescence is a purely optical, non-contact strategy for measuring blood flow and oxygen tension within the cortex. This system builds upon prior work that used structured illumination to overcome traditional limitations of lifetime imaging \cite{Ponticorvo:2010uv}. Spatially patterning excitation light with a DMD allows for the use of a point detector, which offers both high sensitivity and speed for the collection of the phosphorescent signal. The spatial and temporal resolutions of the system are ultimately limited by noise. Regions as small as 0.01 mm$^{2}$ can reliably be excited at a pattern repetition rate of 10 Hz with 200 phosphorescent decays collected and averaged per pattern. Because intensity of the excitation light scales with the size of the target area, larger regions could allow for faster pattern rates at the expense of averaging (e.g. 100 Hz with 20-decay averaging). However, a 10 Hz pattern rate is more than sufficient for visualizing dynamic physiological events such as spreading depolarizations, which have been reported to propagate at a rate of several millimeters per minute \cite{Lauritzen:1994vs}.

@@

\acknowledgments
This study was supported by the National Institutes of Health (EB011556, EB018464, NS078791, NS082518, NS092986) and the American Heart Association (14EIA8970041).
